Random Forests", OAs

### 2023-11-20 src/la_RF.Rmd
CAST

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)
library(caret)
library(randomForest)

@@

#########CHECK 1:9################

dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

@@

##############ADD EAST, 4 REGIONS IN TOTAL###########

# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("E12000007", "E12000008", "E12000009" )
df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))

@@

### train

@@

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "rf", # xgbTree "rpart", # rf lm
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

>>>>

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)
library(caret)
library(randomForest)
library(CAST)

@@

# dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
#   rename(nearest.city.LAD21CD = LAD21CD) %>% 
#   dplyr::select(ladcd, nearest.city.LAD21CD)

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

@@

# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("E12000006", "E12000007", "E12000008", "E12000009" )
df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))

@@

@@

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "ranger", 
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

## RF spatial with CAST

df <- df %>% filter(year > 1996)

#df <- df %>% dplyr::select(-n.lag)
```

### train the model in all data

```{r}
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

start.time <- Sys.time()
model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
#                                                         "year")], 
#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
#                     trControl = tc,
#                     method = "ranger",
#                     #na.action = na.omit,
#                     #importance = "impurity",
#                     metric="Rsquared",
#                     verbose=FALSE)
# ffsmodel_LLO
# ffsmodel_LLO$selectedvars
# plot_ffs(ffsmodel_LLO)
```

```{r}

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

# train in every region
start.time <- Sys.time()

regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN21CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all)
}

rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  theme(legend.position = "none")
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### 2023-11-20 src/oa_RF.Rmd
CAST

@@

```{r}
out.path <- paste0(path, "/data/temp/oa_rf.RData")
# save.image(file = out.path)
start.time <- Sys.time()
load(out.path)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

@@

#stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

@@

varimp_mars <- varImp(m)
plot(varimp_mars, main="Variable Importance")
```

@@

```{r}
pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
length(pred)

df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
postResample(pred = pred$value, obs = pred$n) ## 0.99
```

@@

# run the model for every 2 years
start.time <- Sys.time()
years <- 1997:2012

for (t in years){
  tplus1 <- t + 1
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "rf", #"rpart", # rf lm
                     na.action = na.omit,
                     preProc = c("center", "scale"),
                     importance = TRUE)
}

@@

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.all\\d"))

@@

# use the model trainned in year t to predict io of year t + 1
years <- (1998:2011)
pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
for (t in years){
  t_plus_1 <- t + 1
  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
  pred <- as.data.frame(pred)
  rownames(pred) <- c()
  pred.by.year.all <- cbind(pred.by.year.all, pred)
}

@@

# drop 2000 and select variables form total
df.no1997_1 <- df %>% 
  filter(year!=1996 & year!=1997 & year!=1998) %>% 
  dplyr::select(ladcd, year, n) # CHANGE n WITH growth

# column bind prediction and data  
pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE

@@

>>>>

sapply(df, function(x) sum(is.na(x)))

# out.path <- paste0(path, "/data/temp/oa_rf.RData")
# save.image(file = out.path)
# load(out.path)

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
#                         "Leicester", "nearest.city.oa11cd")) %>% 
#   write_csv(file = out.path)
```

@@

```{r}
# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)

@@

#stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #1h

@@

varimp_mars <- varImp(m.test)
plot(varimp_mars, main="Variable Importance")
```

@@

```{r}
pred <- predict(m.test, df %>% ungroup() %>% filter(RGN11CD=="E12000006"), na.action = na.pass)
length(pred)

df %>% ungroup() %>% filter(RGN11CD=="E12000006") %>% filter(if_any(everything(), is.na))

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN11CD=="E12000006"))
postResample(pred = pred$value, obs = pred$n) ## 0.99
```

@@

detectCores()
cl <- makePSOCKcluster(6)#14
registerDoParallel(cl)

# run the model for every 2 years
start.time <- Sys.time()
years <- 2001:2012 #1997

for (t in years){

  tplus1 <- t + 1
  set.seed(71)
  print(t)

  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "ranger", 
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = 'impurity') #TRUE

  #assign(paste0("model.all",as.character(t + 1)), model.all)
  file <- paste0(path, "/data/temp/oa_rf/timeseries", as.character(t + 1), ".RData")
  save(model.all, file = file)
  rm(model.all)
}

@@

```{r}
# create a list of the above model outputs

load(file = (paste0(path, "/data/temp/oa_rf/timeseries2001.RData")))

filenames <- list.files(paste0(path, "/data/temp/oa_rf"), pattern = "*.RData", full.names = T)

t <- 1997
for (i in filenames){
  load(file = i)
  t <- t + 1
  assign(paste0("timeseries",t), model.all)  
}

fit.model.all <- mget(ls(pattern = "^timeseries\\d"))
rm(list = ls(pattern = "^timeseries\\d"))

@@

# use the model trainned in year t to predict io of year t + 1
years <- (1998:2011)
pred.by.year.all <- data.frame(matrix(NA,nrow=227759, ncol = 0))
for (t in years){
  t_plus_1 <- t + 1
  pred <- predict(fit.model.all[names(fit.model.all)==paste0("timeseries",t)], df[df$year==t_plus_1,])
  pred <- as.data.frame(pred)
  rownames(pred) <- c()
  pred.by.year.all <- cbind(pred.by.year.all, pred)
}

@@

# drop 2000 and select variables form total
df.no1997_1 <- df %>% 
  filter(year!=1996 & year!=1997 & year!=1998) %>% 
  dplyr::select(oa11cd, year, n) # CHANGE n WITH growth

# column bind prediction and data  
pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, oa11cd)) #DOUBLE CHECK ARRANGE

@@

## RF spatial

```{r}
df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% dplyr::select(-n.lag)

```{r}
#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

# run the model for every 2 years
start.time <- Sys.time()
regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "ranger",
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = 'impurity') #TRUE
  assign(paste0("model.all",as.character(i)), model.all)
}

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# train on one region, test on all the rest

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)

    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

rf.year.all.metrics %>% 
  as.data.frame() %>%
  add_rownames(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared)
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### test plots

```{r}
#plots
ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```
